Scaling Automated Theory Exploration

### 2015-08-06 AI4FM/abstract.tex
Edits

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Chris Warburton},
            pdftitle={Scaling Automated Theory Exploration},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

By removing user interaction, automated TE systems require an algorithm
for deciding whether a theorem is ``interesting''. In existing systems,
this is intimately connected to the choice of search algorithm, to
ensure tractability. For example, \textsc{IsaCoSy} \cite{johansson2009isacosy}
discovers equations, which are defined as
``interesting'' if they cannot be simplified using previously discovered
equations. The intuition for such criteria is to avoid special cases of
known theorems, such as \(0 + 0 = 0\), \(0 + 1 = 1\), etc. when we
already know \(\forall x. 0 + x = x\). To work effectively, this
requires general forms to be found \emph{before} special cases.

TE is a ``bottom-up'' process: new theorems are derived from given
knowledge and previous theorems. In particular, there is no special
``goal'' theorem(s) as found in interactive theorem proving; all we have
is our (proxy) ``interesting'' criterion, and the definitions of our
theory. This makes theory exploration a \emph{combinatorial
optimisation} problem, which is well studied in Artificial Intelligence
and Machine Learning. Existing TE systems rely on complete, brute-force
search techniques which do not scale to theories of a significant size;
these could be replaced by a variety of faster, approximate algorithms,
eg. those surveyed in \cite{blum2011hybrid}.

Whilst efficient on a small scale, where we expect the chosen inputs to
be related, we cannot expect an effective output criterion to coincide
so readily with search algorithm and output criterion to be effective
for large inputs containing many irrelevant connections. umbers of also
consider data mining approaches, which infer general rules \emph{after}
enumerating some number of special cases. A variety of
``interestingness'' measures are used in domains such as data mining
\cite{geng2006interestingness}, which may be adapted to a TE setting. These
include objective measures, eg. based on information theory, and
subjective measures which ask for the user's opinion. We can avoid
asking users directly, by treating existing proof libraries and test
suites as representative examples of what is interesting, and hence
construct an oracle; such oracles have already been used for
precision-recall experiments to evaluate the performance of
\textsc{HipSpec} \cite{claessen2013automating}.

@@

Automated theory exploration has been applied to libraries in Isabelle
and Haskell, although we focus on the latter as its implementations are
the most mature (demonstrated by \textsc{Hipster}, which explores Isabelle by
translating code to Haskell first). Haskell is interesting to target,
since its emphasis on pure functions and algebraic datatypes make it common for
functions to obey equational properties; although Haskell is not a theorem prover,
and hence cannot represent such equations . Also, the widespread use of
recursion and higher-order functions makes automated deduction non-trivial
effort is spent discovering and proving these properties compared to
proof-oriented systems like Isabelle.

>>>>

\documentclass[]{eceasst}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
% \ifxetex
%   \usepackage[setpagesize=false, % page size defined by xetex
%               unicode=false, % unicode breaks when used with xetex
%               xetex]{hyperref}
% \else
%   \usepackage[unicode=true]{hyperref}
% \fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Chris Warburton},
            pdftitle={Scaling Automated Theory Exploration},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

By removing user interaction, automated TE systems require an algorithm
for deciding whether to inform the user of a particular theorem or not; we refer
to this search criterion as the theorem being ``interesting''.

In existing systems, this criterion is intimately connected to the
search algorithm; this coupling improves efficiency, even for brute-force
search. As an example, \textsc{IsaCoSy} \cite{johansson2009isacosy} discovers
equations, which are defined as ``interesting'' if they cannot be simplified
using previously discovered equations. The intuition for such criteria is to
avoid special cases of known theorems, such as \(0 + 0 = 0\), \(0 + 1 = 1\),
etc. when we already know \(\forall x. 0 + x = x\). To work effectively, this
requires general forms to be found \emph{before} special cases.

More generally, automated TE is an \emph{optimisation problem}; an area which
has been studied extensively in the fields of Artificial Intelligence and
Machine Learning. For a more thorough treatment, \cite{geng2006interestingness}
surveys recent work on search algorithms and \cite{blum2011hybrid} surveys
interestingness criteria in the context of data mining.

@@

Automated theory exploration has been applied to libraries in Isabelle
and Haskell, although we focus on the latter as its implementations are
the most mature (demonstrated by \textsc{Hipster}, which explores Isabelle by
translating code to Haskell first). Haskell is interesting to target,
since its emphasis on pure functions and algebraic datatypes make equational
properties common; however, Haskell is not a theorem prover, and hence it cannot
represent such as universally quantified equations. Also, the widespread use of
recursion and higher-order functions makes automated deduction non-trivial
effort is spent discovering and proving these properties compared to
proof-oriented systems like Isabelle.
